Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-11-04 robdml.tex
para

In this paper, we first define the ``\emph{hardness}'' of a sample triplet as
the difference between the anchor-positive distance and the anchor-negative
distance.
%
Then, Hardness Manipulation (HM) is proposed to adversarially perturb a given
sample triplet and increase its hardness into a specified \emph{destination}
hardness level.
%
HM is implemented as an objective to minimize the L-$2$ norm of the thresholded
difference between the hardness of the given sample triplet and the specified
\emph{destination} hardness.
%
Mathematically, when the objective is optimized using Projected Gradient
Descent~\cite{madry}, the gradient of HM objective with respect to the
adversarial perturbation is the same as that of directly \emph{maximizing} the
triplet loss.
%
Thus, the optimization process of HM objective can be interpreted as a direct
\emph{maximization} process of the loss (indicating a direct strong adversary)
which stops halfway at the specified \emph{destination} hardness level.

\oo{regular and max can be expressed as two boundary cases of HM}

>>>>

In this paper, we first define the ``\emph{hardness}'' of a sample triplet as
the difference between the anchor-positive distance and anchor-negative
distance.
%
Then, Hardness Manipulation (HM) is proposed to adversarially perturb a given
sample triplet and increase its hardness into a specified \emph{destination}
hardness level for adversarial training.
%
The objective of HM is to minimize the L-$2$ norm of the thresholded difference
between the hardness of the given sample triplet and the specified
\emph{destination} hardness.
%
HM is flexible as regular training and min-max adversarial training
can be expressed as its boundary cases, as shown in \cref{fig:hmflexible}.
%
Mathematically, when the HM objective is optimized using Projected Gradient
Descent~\cite{madry}, the sign of its gradient with respect to the adversarial
perturbation is the same as that of directly \emph{maximizing} the loss.
%
Thus, the optimization of HM objective can be interpreted as a direct and
efficient \emph{maximization} process of the loss which stops halfway at the
specified \emph{destination} hardness level, \ie, the ``intermediate'' point.
